---
title: "Week 4"
author: "Martin Aragoneses"
date: "July 28, 2015"
output: pdf_document
---

## The structure `str()` function 

Alternative to summary()... whats in this object?

```{r}
str(airquality)
s <- split(airquality, airquality$Month)
str(s)
```

## Simple simulation

Generating random numbers
```{r}
# r for random number generation
# d for density
# p for cumulative distribution (CDF)
# q for quantile distribution
str(rnorm)
str(dnorm)
str(pnorm)
str(qnorm)
set.seed(1)
rnorm(5)
rnorm(5)
set.seed(1) # reset seed
rnorm(5)

## Cummulative distribution
# Pr(x<=2) if rate (mean, lambda) is 2
ppois(2, 2)
# Pr(x<=4) if rate (mean, lambda) is 2
ppois(4, 2)
# Pr(x<=6) if rate (mean, lambda) is 2
ppois(6,2)

```

## Simulation from linear model

```{r}
# y = b_0 + b_1*x + e
# e ~ N(0,2^2), x ~ N(0, 1^2), (b_0, b_1) = (0.5, 2)
# 100 observations

set.seed(2)
x <- rnorm(100)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2*x + e
summary(y)
plot(x, y)

```

If x is some treatment, we can generate it from binary data

```{r}
set.seed(2)
x <- rbinom(100, 1, 0.5)
e <- rnorm(100, 0, 2)
y <- 0.5 + 2*x + e
summary(y)
plot(x, y)
```

With count y (instead of continuous), the error is not going to be normal, we can use Poisson
```{r}
## Y ~ Poisson(mu) ## Poisson with mean mu
## log(mu) = b_0 + b_1*x
## b_0 = 0.5, b_1 = 0.3
set.seed(3)
x <- rnorm(100)
mu <- exp(0.5 + 0.3*x)

y <- rpois(100, exp(log(mu)))
summary(y) 
plot(x, y)
```

## Simulating random sampling

Draws randomly from a vector of objects

```{r}
str(sample)
set.seed(4)
sample(1:10, 4)
sample(1:10, 4)
set.seed(4)
sample(1:10, 4)

sample(letters, 5)

sample(1:10) ## permutation
sample(1:10, replace = TRUE) ## with replacement\
```

## R Profiler
```{r}
## system.time()  how long does it take to evaluate an expression
## summaryRprof() summarizes the output form Rprof()
## use the by.self to subsstract the lower level functions

```
